---
title: <center><strong> "Classification and Regression Trees"</strong></center>
author: <center><strong>[Shelmith Nyagathiri Kariuki](https://github.com/Shelmith-Kariuki)</strong></center>
date: <center><strong>"August 8, 2019"</strong></center>
output:
  html_document:
    toc_depth: '6'
    toc: yes
---

The data used in this exercise has been obtained from the Financial Inclusion Competition hosted on [Zindi.africa](http://zindi.africa/competitions/financial-inclusion-in-africa)


```{r setup, include=FALSE}


knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

## Install the libraries required

```{r}
## create a vector of packages to be installed
pkgs <- c("dplyr","caret","randomForest")

## Check if there are packages you want to load, that are not already installed. 
miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]] 

## Installing the missing packages
if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}
## Loading all the packages
invisible(lapply(pkgs,library,character.only=TRUE))

## Remove the objects that are no longer required
rm(miss_pkgs)
rm(pkgs)
```

## Read in the datasets

```{r}

data_path <- "/home/shelmith/Documents/Github/zuka/Datasets/FinancialInclusion"

train_dat <- read.csv(paste(data_path, "Train_v2.csv",sep="/"))

test_dat <- read.csv(paste(data_path, "Test_v2.csv",sep="/"))

```

## Check for missing values

```{r, comment=NA}

## train data
apply(train_dat,2,function(x) sum(is.na(x)))

## test data
apply(test_dat,2,function(x) sum(is.na(x)))

```


## Drop the variables that we will not need

```{r, comment=NA}

## train data
train <- train_dat %>% 
  select(-uniqueid, -year)

## test data
test <- test_dat %>% 
  select(-uniqueid, -year)

```

## Carry out anova tests to check whether there is a significant difference in ownership of a bank account among the different categories

```{r, comment=NA}

## Convert the bank_account into 0 and 1

train$bank_account <-ifelse(train$bank_account == "Yes",1,0)

## country
summary(aov(bank_account~country, data = train))
TukeyHSD(aov(bank_account~country, data = train))
##summary(lm(bank_account~country, data = train))


## location_type
t.test(bank_account~location_type, data = train) ## Those in urban areas are more likely to own a bank account, as opposed to those in the rural areas 

## cellphone_access
t.test(bank_account~cellphone_access, data = train) ## Those with cellphones are more likely to own a bank account, as opposed to those without


## household_size
cor.test(train$household_size,train$bank_account, method = "spearm", alternative = "g")


## age_of_respondent
cor.test(train$age_of_respondent,train$bank_account, method = "spearm", alternative = "g")

## gender_of_respondent
t.test(bank_account~gender_of_respondent, data = train) ## Males are more likely to own a bank account, as opposed to females.

## relationship_with_head
summary(aov(bank_account~relationship_with_head, data = train))
TukeyHSD(aov(bank_account~relationship_with_head, data = train))

## marital_status

summary(aov(bank_account~marital_status, data = train))
TukeyHSD(aov(bank_account~marital_status, data = train))


## education_level
## No formal education, Other/Dont know/RTA -None

train$education_level <- as.factor(ifelse(train$education_level %in% grep("formal|No",train$education_level,value = T, ignore.case = T),"None",as.character(train$education_level)))
                         
 test$education_level <- as.factor(ifelse(test$education_level %in% grep("formal|No",test$education_level,value = T, ignore.case = T),"None",as.character(test$education_level)))

 
summary(aov(bank_account~education_level, data = train))
TukeyHSD(aov(bank_account~education_level, data = train))

## job_type

## Collapse this variable into fewer categorical variables
# Formally employed Private,Formally employed Government - Employed 
# Self employed
# Remittance Dependent,Government Dependent, Other Income - Other
# Dont Know/Refuse to answer, No Income - None

## train data
train$job_type <- as.factor(ifelse(train$job_type %in% grep("Formally",train$job_type,value = T, ignore.case = T),"Employed",
          ifelse(train$job_type %in% grep("Dependent|Other",train$job_type,value = T, ignore.case = T),"Other",
        ifelse(train$job_type %in% grep("Dont Know/Refuse to answer|No Income",train$job_type,value = T, ignore.case = T),"None",as.character(train$job_type)))))


## test data
test$job_type <- as.factor(ifelse(test$job_type %in% grep("Formally",test$job_type,value = T, 
                                                ignore.case = T),"Employed",
          ifelse(test$job_type %in% grep("Dependent|Other",test$job_type,value = T, ignore.case = T),"Other",
        ifelse(test$job_type %in% grep("Dont Know/Refuse to answer|No Income",test$job_type,value = T, ignore.case = T),"None",as.character(test$job_type)))))

summary(aov(bank_account~job_type, data = train))
TukeyHSD(aov(bank_account~job_type, data = train))

```

## Decision Tree

```{r, comment=NA}
trctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
set.seed(3333)
dec.tree = train(bank_account ~ ., 
                  data=train, 
                  method="rpart",
                  trControl = trctrl)
dec.tree
```

```{r, comment=NA}

train_pred <- predict(dec.tree, newdata = train)
train_pred<- ifelse(train_pred<0.45,0,1)
#confusionMatrix(train_pred, train$bank_account)

confusionMatrix(table(train_pred, train$bank_account)) 
```

```{r, comment=NA}

test_pred <- predict(dec.tree, newdata = test)
test_pred<- ifelse(test_pred<0.45,0,1)

test_dat$bank_account = test_pred

test_sub <- test_dat %>% 
  select(uniqueid, bank_account, country)

test_sub$uniqueid <-paste(test_sub$uniqueid,"x",test_sub$country)

## Read in submission file 

sub_file <- read.csv(paste(data_path, "SubmissionFile.csv",sep="/"))
sub_file$bank_account <- NULL
test_sub <-inner_join(sub_file, test_sub,by = "uniqueid")

write.csv(test_sub, paste(data_path, "test_sub1.csv",sep="/"),row.names = FALSE) #Leaderboard position 69, with a value of 0.137418203
```

## Bagging: Random Forest

```{r, comment=NA}
apply(train,2,function(x) sum(is.na(x)))

# Fitting model
fit <- randomForest(bank_account ~ ., data=train ,ntree=500)
summary(fit)

#Predict Output 
predicted= predict(fit,train)


train_pred <- predict(fit, newdata = train)
train_pred<- ifelse(train_pred<0.45,0,1)
#confusionMatrix(train_pred, train$bank_account)

confusionMatrix(table(train_pred, train$bank_account)) 


```


```{r, comment=NA}


test_pred <- predict(fit, newdata = test)
test_pred<- ifelse(test_pred<0.45,0,1)

test_dat$bank_account = test_pred

test_sub <- test_dat %>% 
  select(uniqueid, bank_account, country)

#test_sub$uniqueid <-paste(test_sub$uniqueid,"x",test_sub$country)

## Read in submission file 

sub_file <- read.csv(paste(data_path, "SubmissionFile.csv",sep="/"))
sub_file$bank_account <- NULL
sub_file$X.1 <- NULL

names(sub_file) = c("uniqueid","country")
test_sub <-inner_join(sub_file, test_sub,by = c("uniqueid","country"))
test_sub$uniqueid <-paste(test_sub$uniqueid,"x",test_sub$country)
sub_file$country <- NULL


write.csv(test_sub, paste(data_path, "test_sub2.csv",sep="/"),row.names = FALSE) 

```


## Boosting: Gradient Boosting Model (GBM)